---
title: "ETC3250 Assignment 2"
date: "SOLUTION"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE)
```

## Exercises

1. This question is about the normal distribution, and how it relates to the classification rule provided by linear discriminant analysis.

    a. Write down the density function for a bivariate normal distribution ($p=2$), with mean $\mu_k$ and variance $\Sigma$. 
    b. Show that the linear discriminant rule for two groups ($K=2$), $\pi_1=\pi_2$ and $\Sigma_1=\Sigma_2 = \Sigma = \left[\begin{array}{cc} \sigma_1^2 & \rho\sigma_1\sigma_2 \\ \rho\sigma_1\sigma_2 & \sigma_2^2 \end{array}\right]$ where $\rho$ is the population correlation between the two variables, and $\sigma_1, \sigma_2$ are the population standard deviations of the two variables, respectively, is equal to:
*Assign a new observation $x_0$ to group 1 if*
$$x_0'\Sigma^{-1}(\mu_1-\mu_2) > \frac{1}{2}(\mu_1 + \mu_2)'\Sigma^{-1}(\mu_1-\mu_2)$$
    c. By generating a grid of values, draw the boundary between two groups, in the 2D space. Use these values for $\mu_1, \mu_2$ and $\sigma$.
$$\mu_1 = \left[\begin{array}{r} -2 \\ 2 \end{array}\right], ~~~\mu_2 = \left[\begin{array}{r} 2 \\ -2 \end{array}\right]$$
$$\Sigma = \left[\begin{array}{rr} 4 & 3 \\ 3 & 5 \end{array}\right]$$
    d. Write down the rule using these parameter values, and sketch the line corresponding to the 1D discriminant space on the previous plot.

2. This question is about dimension reduction using PCA. We will use data from the world bank, on development indicators for 264 countries. Go to this site  https://databank.worldbank.org/data/source/world-development-indicators/preview/on# to extract a copy of the data yourself. You should select all the countries that they have available, and just year 2017. There are some countries that are not countries, that you should exclude: "Upper middle income", "Pre-demographic dividend", "Post-demographic dividend", "Other small states", "OECD members", "Not classified", "Middle income", "Lower middle income", "Low income", "Low & middle income", "Least developed countries", "UN classification", "Late-demographic dividend", "IDA & IBRD total", "IDA blend", "IDA only", "IDA total", "High income", "Heavily indebted poor countries (HIPC)", "Fragile and conflict affected situations". 

Use the default set of variables that are chosen: 

```{r}
library(readxl)
library(tidyverse)
WDI <- read_xlsx("data/Data_Extract_From_World_Development_Indicators.xlsx", na="..", n_max=14466)
exclude <- c("Upper middle income","Pre-demographic dividend", "Post-demographic dividend", "Other small states", "OECD members", "Not classified", "Middle income", "Lower middle income", "Low income", "Low & middle income", "Least developed countries: UN classification", "Late-demographic dividend", "IBRD only", "IDA & IBRD total", "IDA blend", "IDA only", "IDA total", "High income", "Heavily indebted poor countries (HIPC)", "Fragile and conflict affected situations", "Early-demographic dividend", "Data from database: World Development Indicators")
WDI <- WDI %>% filter(!(`Country Name` %in% exclude))
WDI %>% count(`Series Name`) %>% print(n=60)
WDI %>% count(`Country Name`) %>% print(n=300)
```

You will need to do some cleaning on the data. 

i. Remove the two lines which are missing on the `Series Name`, and rename `2017 [YR2017]` to "value".
ii. Make a look up dictionary mapping `Series Code` to `Series Name`, so that we can do the analysis with the code (shorter), but refer back to the name as needed.
iii. Spread the data into wide form, so that you have variables `Country Name`, `Country Code`, and the 55 `Series Name` variables as columns.
iv. Using the `naniar` package make a missingness heatmap of the data
v. Remove any variable that is missing for more than 100 countries. Then, remove any countries that have missing on more than 2 variables.
vi. Use $k$ nearest neighbours imputation to fill in the missings.
        
```{r}
# Remove rows with missings and rename measured variable to value
WDI <- WDI %>% filter(!is.na(`Series Name`)) %>%
  rename(value = `2017 [YR2017]`)
# Create lookup dictionary
WDI_names <- WDI %>% 
  select(`Series Code`, `Series Name`) %>%
  distinct()
# Make wide form to inspect missing data
library(naniar)
library(impute)
WDI_wide <- WDI %>% 
  select(-`Series Name`) %>%
  spread(`Series Code`, value)
vis_miss(WDI_wide, sort_miss=TRUE, cluster=TRUE)
# Remove variables and countries with too many missings
WDI_varmiss <- WDI %>% filter(is.na(value)) %>% count(`Series Code`, sort=TRUE) %>%
  filter(n < 100) # Missing on 2/3 countries removed
WDI_nomiss <- WDI %>% filter(`Series Code` %in% WDI_varmiss$`Series Code`) # Down to 32 variables
WDI_cntmiss <- WDI_nomiss %>% filter(is.na(value)) %>% count(`Country Code`, sort=TRUE) %>% filter(n<3) # Keep countries with only two missings
WDI_nomiss <- WDI_nomiss %>% filter(`Country Code` %in% WDI_cntmiss$`Country Code`)
# Set up for knn, and impute missings
WDI_nomiss_wide <- WDI_nomiss %>% select(-`Series Name`) %>% 
  spread(`Series Code`, value)
WDI_impute <- impute.knn(as.matrix(WDI_nomiss_wide[,-c(1,2)]))$data
WDI_nomiss_wide[,3:34] <- WDI_impute
```

```{r eval=FALSE}
# This is the code to install the impute package with a knn imputation fn
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("impute", version = "3.8")
```

a. How many countries are in the full data set? How many variables?
b. How many variables are missing on more than 100 countries?
c. How many countries have missing values on more than 2 variables, after the variables in b. have been removed?
d. Compute a principal component analysis for the cleaned and imputed data matrix. Tabulate the proportion of variation explained up to 8 PCs. 
e. Make a scree plot. How many principal components would be suggested? What proportion of variation would be explained by your choice? Please explain your thinking and decisions.

```{r}
WDI_pca <- prcomp(WDI_nomiss_wide[,3:34], retx=TRUE, center=TRUE, scale=TRUE)
WDI_pca_var <- tibble(n=1:length(WDI_pca$sdev), evl=WDI_pca$sdev^2)
ggplot(WDI_pca_var, aes(x=n, y=evl)) + geom_line() +
  xlab("Number of PCs") + ylab("Eigenvalue")
```

f. Make a biplot of the first two principal components, and explain the contributions of the variables, and similarity of the variables.

```{r}
library(gridExtra)
WDI_pca_pcs <- as_tibble(WDI_pca$x[,1:3]) %>%
  mutate(cnt=WDI_nomiss_wide$`Country Code`, 
         name=WDI_nomiss_wide$`Country Name`)
WDI_pca_evc <- as_tibble(WDI_pca$rotation[,1:3]) %>% 
  mutate(origin=rep(0, 32), variable=as.character(1:32),
         varname=rownames(WDI_pca$rotation)) %>%
  mutate(PC1s = PC1*(WDI_pca_var$evl[1]*2), 
         PC2s = PC2*(WDI_pca_var$evl[2]*2),
         PC3s = PC3*(WDI_pca_var$evl[3]*2))
p1 <- ggplot() + 
  geom_segment(data=WDI_pca_evc, aes(x=origin, xend=PC1s, y=origin, yend=PC2s), colour="red") +
  geom_text(data=WDI_pca_evc, aes(x=PC1s, y=PC2s, label=variable), colour="red") +
  geom_point(data=WDI_pca_pcs, aes(x=PC1, y=PC2)) +
  geom_text(data=filter(WDI_pca_pcs, abs(PC1)>5), aes(x=PC1, y=PC2, label=cnt)) +
  geom_text(data=filter(WDI_pca_pcs, abs(PC2)>5), aes(x=PC1, y=PC2, label=cnt)) +
  xlab("PC1") + ylab("PC2") +
  theme(aspect.ratio=1)
p2 <- ggplot() + 
  geom_segment(data=WDI_pca_evc, aes(x=origin, xend=PC1s, y=origin, yend=PC3s), colour="red") +
  geom_text(data=WDI_pca_evc, aes(x=PC1s, y=PC3s, label=variable), colour="red") +
  geom_point(data=WDI_pca_pcs, aes(x=PC1, y=PC3)) +
  geom_text(data=filter(WDI_pca_pcs, abs(PC1)>5), aes(x=PC1, y=PC3, label=cnt)) +
  geom_text(data=filter(WDI_pca_pcs, abs(PC3)>5), aes(x=PC1, y=PC3, label=cnt)) +
  xlab("PC1") + ylab("PC3") +
  theme(aspect.ratio=1)
p3 <- ggplot() + 
  geom_segment(data=WDI_pca_evc, aes(x=origin, xend=PC2s, y=origin, yend=PC3s), colour="red") +
  geom_text(data=WDI_pca_evc, aes(x=PC2s, y=PC3s, label=variable), colour="red") +
  geom_point(data=WDI_pca_pcs, aes(x=PC2, y=PC3)) +
  geom_text(data=filter(WDI_pca_pcs, abs(PC2)>5), aes(x=PC2, y=PC3, label=cnt)) +
  geom_text(data=filter(WDI_pca_pcs, abs(PC3)>5), aes(x=PC2, y=PC3, label=cnt)) +
  xlab("PC2") + ylab("PC3") +
  theme(aspect.ratio=1)
grid.arrange(p1, p2, p3, ncol=3)
```

g. Examine the largest coefficients for the first three principal components. Explain and interpret the first three principal components. 

```{r}
WDI_pca_evc %>% select(variable, varname, PC1) %>% arrange(desc(PC1)) %>% print(n=32)
WDI_pca_evc %>% 
  select(variable, varname, PC1) %>% 
  filter(abs(PC1)>0.2) %>% 
  left_join(WDI_names, by=c("varname"="Series Code")) %>%
  arrange(PC1)
WDI_pca_evc %>% select(variable, varname, PC2) %>% arrange(desc(PC2)) %>% print(n=32)
WDI_pca_evc %>% 
  select(variable, varname, PC2) %>% 
  filter(abs(PC2)>0.2) %>% 
  left_join(WDI_names, by=c("varname"="Series Code")) %>%
  arrange(PC2)
WDI_pca_evc %>% select(variable, varname, PC3) %>% arrange(desc(PC3)) %>% print(n=32)
WDI_pca_evc %>% 
  select(variable, varname, PC3) %>% 
  filter(abs(PC3)>0.2) %>% 
  left_join(WDI_names, by=c("varname"="Series Code")) %>%
  arrange(PC3)
```

*PC1 is an indicator for development, with statistics such as mortality, rural industry, population, contrasting personal income and immunization rates. High values indicate high development (eg Europe & Central Asia, Netherlands, Luxemburg), and low values less development (eg Chad, Sudan, Mali). PC2 is a contrast of trade to income, population and foreign investment. PC3 is a contrast of industrialisation and conservation to population density, rural industry and growth in GDP.*

h. Several countries stand out as outliers on the biplots. Examine the principal component scores for the first three principal components. Identify the countries that are at the extremes of each, and explain what their characteristics are. Make a plot to support your arguments.
    
```{r}
WDI_pca_pcs %>% 
  select(cnt, name, PC1) %>% 
  filter(abs(PC1)>5) %>% 
  arrange(PC1)
ggplot(WDI_nomiss_wide, aes(x=SH.DYN.MORT)) + geom_density(fill="black", alpha=0.3) +
  geom_vline(data=filter(WDI_nomiss_wide, `Country Name` %in% c("Chad", "Europe & Central Asia")), aes(xintercept=SH.DYN.MORT, colour=`Country Name`))
ggplot(WDI_nomiss_wide, aes(x=NY.GNP.PCAP.PP.CD)) + geom_density(fill="black", alpha=0.3) +
  geom_vline(data=filter(WDI_nomiss_wide, `Country Name` %in% c("Chad", "Europe & Central Asia")), aes(xintercept=NY.GNP.PCAP.PP.CD, colour=`Country Name`))
WDI_pca_pcs %>% 
  select(cnt, name, PC2) %>% 
  filter(abs(PC2)>5) %>% 
  arrange(PC2)
ggplot(WDI_nomiss_wide, aes(x=SP.POP.TOTL)) + geom_density(fill="black", alpha=0.3) +
  geom_vline(data=filter(WDI_nomiss_wide, `Country Name` %in% c("Luxembourg", "Europe & Central Asia")), aes(xintercept=SP.POP.TOTL, colour=`Country Name`))
ggplot(WDI_nomiss_wide, aes(x=NE.IMP.GNFS.ZS)) + geom_density(fill="black", alpha=0.3) +
  geom_vline(data=filter(WDI_nomiss_wide, `Country Name` %in% c("Luxembourg", "Europe & Central Asia")), aes(xintercept=NE.IMP.GNFS.ZS, colour=`Country Name`))
WDI_pca_pcs %>% 
  select(cnt, name, PC3) %>% 
  filter(abs(PC3)>3) %>% 
  arrange(PC3)
ggplot(WDI_nomiss_wide, aes(x=EN.POP.DNST)) + geom_density(fill="black", alpha=0.3) +
  geom_vline(data=filter(WDI_nomiss_wide, `Country Name` %in% c("Sierra Leone", "Congo, Rep.")), aes(xintercept=EN.POP.DNST, colour=`Country Name`))
ggplot(WDI_nomiss_wide, aes(x=NV.IND.TOTL.ZS)) + geom_density(fill="black", alpha=0.3) +
  geom_vline(data=filter(WDI_nomiss_wide, `Country Name` %in% c("Sierra Leone", "Congo, Rep.")), aes(xintercept=NV.IND.TOTL.ZS, colour=`Country Name`))
```

*On PC1, Europe & Central Asia have high positive scores and Chad has the lowest negative score. Chad would be considered to be a less developed country, in contrast to Europe & Central Asia. On PC2 Luxembourg has high positive scores, and Europe & Central Asia has low scores. Luxembourg is a tiny powerhouse of trade relative to the big giants of Europe, and even China. Congo has high positive scores on PC3 and Sierra Leone has low scores. Congo has relatively low population density, and a large proportion of GDP due to industry, and looks to be an African powerhouse. This is a bit surprising, but reading about Congo (on wikipedia) indicates that it is about to boom as one of the few cobalt producers in the world, essential for lithium batteries in electric vehicles. They say its the "Saudia Arabia of the electric vehicle age".*

i. Use bootstrap, resampling the countries to produce 95% confidence intervals for the coefficients on PC1. Revise your interpretation of PC1 based on significance of the coefficients. 
    
```{r}
library(boot)
library(forcats)
compute_PC <- function(data, index, k=1) {
  pc <- prcomp(data[index,], center=TRUE, scale=TRUE)$rotation[,k]
  if (sign(pc[1]) < 0)
    pc <- -pc
  return(pc)
}
# Make sure the first coefficient of the PC is positive.
PC1_boot <- boot(data=WDI_nomiss_wide[,3:34], compute_PC, R=1000)
colnames(PC1_boot$t) <- colnames(WDI_nomiss_wide[,3:34])
PC1_boot_t0 <- tibble(var=factor(names(PC1_boot$t0), levels=names(PC1_boot$t0)), t0=PC1_boot$t0)
PC1_boot_ci <- as_tibble(PC1_boot$t) %>%
  gather(var, coef) %>% 
  mutate(var = factor(var, levels=levels(PC1_boot_t0$var))) %>%
  group_by(var) %>%
  summarise(q2.5 = quantile(coef, 0.025), 
            q5 = median(coef),
            q97.5 = quantile(coef, 0.975)) %>%
  left_join(PC1_boot_t0) 
PC1_boot_ci %>% arrange(t0) %>% print(n=32)
ggplot(PC1_boot_ci, aes(x=fct_reorder(var, q2.5), y=t0)) + 
  geom_hline(yintercept=0, colour="white", size=2) +
  geom_point() + 
  geom_errorbar(aes(ymin=q2.5, ymax=q97.5)) +
  coord_flip()
PC2_boot <- boot(data=WDI_nomiss_wide[,3:34], compute_PC, k=2, R=1000)
colnames(PC2_boot$t) <- colnames(WDI_nomiss_wide[,3:34])
PC2_boot_t0 <- tibble(var=factor(names(PC2_boot$t0), levels=names(PC2_boot$t0)), t0=PC2_boot$t0)
PC2_boot_ci <- as_tibble(PC2_boot$t) %>%
  gather(var, coef) %>% 
  mutate(var = factor(var, levels=levels(PC2_boot_t0$var))) %>%
  group_by(var) %>%
  summarise(q2.5 = quantile(coef, 0.025), 
            q5 = median(coef),
            q97.5 = quantile(coef, 0.975)) %>%
  left_join(PC2_boot_t0) 
PC2_boot_ci %>% arrange(t0) %>% print(n=32)
PC3_boot <- boot(data=WDI_nomiss_wide[,3:34], compute_PC, k=3, R=1000)
colnames(PC3_boot$t) <- colnames(WDI_nomiss_wide[,3:34])
PC3_boot_t0 <- tibble(var=factor(names(PC3_boot$t0), levels=names(PC3_boot$t0)), t0=PC3_boot$t0)
PC3_boot_ci <- as_tibble(PC3_boot$t) %>%
  gather(var, coef) %>% 
  mutate(var = factor(var, levels=levels(PC3_boot_t0$var))) %>%
  group_by(var) %>%
  summarise(q2.5 = quantile(coef, 0.025), 
            q5 = median(coef),
            q97.5 = quantile(coef, 0.975)) %>%
  left_join(PC3_boot_t0) 
PC3_boot_ci %>% arrange(t0) %>% print(n=32)
```

*Only population and size of land area have coefficients that are significantly different from 0. Thus PC1 is only summarising the physical and human size of the country, with positive values indicating a large and highly populated country. The upshot of doing the accuracy for the first three PCs is that the overall result is unreliable. The outliers, which are composite countries may be affecting the result, and it would be better to remove these, and re-do PCA. Similarly, removing several of the variables, like land area and population, which are not conceivably related to development, may help produce more reliable results.*

j. True or false. *Principal* has the same meaning as *principle*.