---
title: "ETC3250 2019 - Lab 7"
author: "Dianne Cook"
date: "Week 7"
output:
  html_document: default
---

```{r, echo = FALSE, message = FALSE, warning = FALSE, warning = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  error = FALSE, 
  collapse = TRUE,
  comment = "#",
  fig.height = 4,
  fig.width = 6,
  fig.align = "center",
  cache = FALSE
)
```


# Class discussion 

For the olive oil data set, the classification tree will use just one variable for its model. It splits on linoleic acid as shown. 

```{r out.width="70%"}
library(tidyverse)
library(rpart)
olive <- read_csv("http://www.ggobi.org/book/data/olive.csv") %>%
  rename(name=X1) %>%
  filter(region != 1) %>%
  dplyr::select(-name, -area) %>%
  mutate(region = factor(region))
rpart(region~., data=olive)
ggplot(olive, aes(x=linoleic, y=region, colour=region)) + 
  geom_point(alpha=0.5) + scale_colour_brewer("", palette="Dark2") +
  theme(legend.position="none") +
  geom_vline(xintercept=1053.5)
```

*Question 1: There is no gap. Do you think this might be a problem with future data? Why?*

If we look at some other variables, oleic and arachidic acid, in relation to linoleic. 

```{r out.width="100%", fig.width=8}
library(gridExtra)
p1 <- ggplot(olive, aes(x=linoleic, y=oleic, colour=region)) + 
  geom_point() + scale_colour_brewer("", palette="Dark2") +
  theme(legend.position="none", aspect.ratio=1) 
p2 <- ggplot(olive, aes(x=linoleic, y=arachidic, colour=region)) + 
  geom_point() + scale_colour_brewer("", palette="Dark2") +
  theme(legend.position="none", aspect.ratio=1) 
grid.arrange(p1, p2, ncol=2)
```

*Question 2: If you got to choose two variables for splitting the two groups, which would you choose, oleic or arachidic, in association with linoleic?*

*Question 3: Suppose you work with linoleic and arachidic. Would quadratic discriminant analysis produce a better separation than the tree? Argue your viewpoint.*

*Question 4: Find a linear combination of linoleic and arachidic, and create a new variable to pass to the tree. Re-fit the tree with this variable instead of the original two. What does the model look like now? Is this better than the original tree?*

*Question 5: In general, why is it often important to create new variables (feature engineering) when building models?*

# Activities

If you need to use the lab time to coordinate the team submission for assignment 3, please do.


